---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r message=FALSE, warning=FALSE}

library(ggplot2)
library(dplyr)
library(plyr)

```

## Loading and preprocessing the data
```{r}
activity.data <- read.csv("activity.csv",header = TRUE, na.strings = "NA")


```


## What is mean total number of steps taken per day?

### Calculate the mean and median of total number of steps taken per day
```{r}

sum.steps <- ddply(activity.data,.(date),summarize,steps=sum(steps))
hist(sum.steps$steps,main ="Frequency of steps",xlab = "Steps")

mean.data <- ddply(activity.data,.(date),summarize,steps=mean(steps))
mean.data

median.data <- ddply(activity.data,.(date),summarize,steps=median(steps))
median.data
```


## What is the average daily activity pattern?
###Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
mean.data<-aggregate(steps ~ interval,data=activity.data,FUN=mean,na.rm=TRUE)


ggplot(mean.data,aes(interval,steps))+geom_line()+xlab("5-Minute Intervals")+ylab("Average Number of Steps per Interval")

```

###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
mean.data[which.max(mean.data$steps),][1]

```

## Imputing missing values
###Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
total.NAs <- sum(is.na(activity.data$steps))
```

Total number of rows with NAs = `r total.NAs`

###Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

####Simple approach here is to use average number of steps for the interval that is missing

```{r}
new.activity.data <- activity.data
new.activity.data[is.na(new.activity.data$steps) == TRUE,]$steps =mean.data$steps[match(mean.data$interval,new.activity.data$interval)]

```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
new.activity.data$weekend <- weekdays.Date(as.Date(new.activity.data$date))%in%c("Saturday","Sunday")
new.activity.data$daytype <- as.factor(ifelse(new.activity.data$weekend == TRUE,"Weekend","Weekday"))

stepsbyInterval <- ddply(new.activity.data,.(interval,daytype),summarize,total=sum(steps))

ggplot(stepsbyInterval, aes(x=interval, y=total, group=daytype, colour=daytype)) +
   geom_line() + geom_point() +
   facet_grid(daytype ~ ., scales="free_y")


```